---
title: "Genotype x genotype interactions in host manipulation - Analyses"
author: "Dan Benesh"
output: github_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)
theme_set(new = theme_bw())
```

# Experiment Background

Many parasites manipulate the behavior of their hosts in ways that seem beneficial. Little is known about genetic variation in this phenotype. Are parasites genetically variable in how much they manipulate host behavior? Do different host genotypes resist manipulation better than others? I infected 5 host strains with 5 parasite strains to quantify the variability in host manipulation due to parasite genes, host genes, and their interactions. The host-parasite system I used was the tapeworm *Schistocephalus solidus* in its copepod first intermediate host.

```{r importdata}
# import data - averaged within recordings
bd <- read.csv(file = "../data/behav_manutracked_reduced_dataset.csv", header = TRUE) # manual data
bdr <- read.csv(file = "../data/behav_autotracked_reduced_dataset.csv", header = TRUE) # auto-tracked data
infd <- read.csv(file = "../data/GxG_inf.csv", header = TRUE, fileEncoding = "UTF-8-BOM") #excel attached BOM to csv
```

```{r}
# take average worm size
infd <- mutate(infd, proc_size = (proc_size1 + proc_size2)/2)
```
```{r}
# add useful variables to manual data
bd <- mutate(bd, speed_per_sec = tot_dist/frames_tracked * 0.5,
       cop_name = substring(fname, 1, 5),
       day_char = substring(fname, 7, 8),
       rec_time = factor(group, levels = c('before the drop', 'after the drop')))%>%
  select(-group)

# combine with inf data
bd <- left_join(bd, select(infd, -proc_size1, -proc_size2, -exposed, -tank, -cerc_multinom, -remarks),
                by = 'cop_name')
# better labels to trts
bd <- mutate(bd, trt = factor(trt, levels = c('unexposed', 'uninfected', 'infected')),
       trt = factor(trt, labels = c('unexposed, control', 'exposed, uninfected', 'infected')),
       parasite_fam = factor(parasite_fam, labels = c('A', 'B', 'C', 'D', 'E')))
```
```{r}
# add useful variables to auto data
bdr <- mutate(bdr, speed_per_sec = tot_dist/frames_tracked * 2,
       cop_name = substring(fname, 1, 5),
       day_char = substring(fname, 7, 8),
       rec_time = factor(group, levels = c('before the drop', 'after the drop')))%>%
  select(-group)

# combine with inf data
bdr <- left_join(bdr, select(infd, -proc_size1, -proc_size2, -exposed, -tank, -cerc_multinom, -remarks),
                by = 'cop_name')
# better labels to trts
bdr <- mutate(bdr, trt = factor(trt, levels = c('unexposed', 'uninfected', 'infected')),
       trt = factor(trt, labels = c('unexposed, control', 'exposed, uninfected', 'infected')),
              parasite_fam = factor(parasite_fam, labels = c('A', 'B', 'C', 'D', 'E')))
```

Copepod movement was video recorded for two minutes. After one minute, the well plate containing the copepods was dropped in a standardized way to simulate a predator attack and 'frighten' the copepods. Each copepod was recorded multiple times throughout the course of parasite development (from uninfective to infective larva).

I explored different behavioral variables that could be extracted from these recordings [here](GxG_01_define_responses.md). I decided to focus on 'activity', which was the average speed copepods moved before and after the drop (it could be equivalently expressed as the distance moved). Activity incorporates both the frequency and magnitude of copepod movements. And though I found that the difference between infected and uninfected copepods was mainly determined by [movement frequency](GxG_01_define_responses.md#proportion-of-time-moving-vs-hop-distance), I decided there was little to gain by breaking this variable down into its component parts.

In that [notebook](GxG_01_define_responses.md), the focus was on variation within the 2-minute recordings. There is another temporal scale in this experiment, behavior over the days of the parasite's development. Treatment differences also need to be considered at this level, because infected copepods may behave differently when harboring infective vs uninfective worms. However, before looking at treatment differences, the focus on behavior from one day to the next brings up the issue that some copepods died during the experiment. Should we filter these out? Or can we retain these data, because the dead copepods did not behave unusually before dying?

# Filter out dead copepods?

First of all, out of `r length(unique(bdr$cop_name))` total copepods with behavioral recordings, how many died during the experiment?

```{r}
x<-select(bdr, cop_name, dead)%>%distinct()
sum(x$dead)
```

That would be a fairly large number of copepods to exclude. Let's explore whether that is necessary. As a start, we can compare the behavior of copepods that died during the experiment with those that did not.

```{r}
bdr_avg <- filter(bdr, !is.na(speed_per_sec))%>%
  group_by(rec_time, dead, day_char)%>%
  summarize(speed_m = mean(speed_per_sec),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))

ggplot(bdr_avg, aes(x = as.numeric(day_char), y = speed_m, shape = factor(dead))) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lci, ymax = uci), 
                position = position_dodge(width = 0.5)) +
  geom_line(aes(linetype = factor(dead))) +
  facet_grid(~rec_time) +
  labs(title = 'auto tracking', x = 'days post infection', y = 'average speed', shape = 'Dead', linetype = 'Dead')
```

The dead ones are less active on average, which seems realistic. However, treatment differences are lumped together in the above plot. Infecteds are less active than uninfecteds, and if they are also more likely to die, then that could drive the difference. However, the proportion of copepods that died was rather comparable across treatments.

```{r}
deads <- select(bdr, cop_name, trt, dead)%>%
  distinct()%>%
  mutate(dead = factor(dead, labels = c('survived', 'died')))%>%
  group_by(trt, dead)%>%
  summarize(n = n())%>%
  spread(dead, n)%>%
  mutate(prop_dead = died/(survived + died))
deads
```

Consequently, it is not surprising, that when we make the same plot, but separate the treatments, we get a similar pattern. The copepods that died exhibited somewhat lower activity levels, particularly later in the experiment.

```{r}
bdr_avg <- filter(bdr, !is.na(speed_per_sec))%>%
  group_by(rec_time, dead, trt, day_char)%>%
  summarize(speed_m = mean(speed_per_sec),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))

ggplot(bdr_avg, aes(x = as.numeric(day_char), y = speed_m, color = trt, shape = factor(dead))) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lci, ymax = uci), 
                position = position_dodge(width = 0.5)) +
  geom_line(aes(linetype = factor(dead))) +
  facet_grid(trt~rec_time) +
  labs(title = 'auto tracking - 95% CI', color = 'Treatment', 
       x = 'days post infection', y = 'average speed', 
       linetype = 'Dead', shape = 'Dead')
```

Is this reason enough to toss data from dead copepods? Copepods died at different times during the experiment, so maybe their behavior only conspicuously changes right before death. Let's plot the behavioral profiles of individual copepods that died during the experiment. Activity across the experiment is shown for individual copepods (lines connect values for individuals). The vertical panels separate copepods by their time of death.

```{r}
ggplot(filter(bdr, dead == 1, days_surv != 9),
       aes(x = as.numeric(day_char), y = speed_per_sec)) + 
  geom_line(aes(group = cop_name), alpha = 0.2) +
  facet_grid(days_surv~rec_time) +
  scale_x_continuous(breaks = c(5,7,9,11,13,15,17,19,21)) +
  labs(title = 'auto tracking', x = 'days post infection', y = 'speed')
```

If copepods move less as they near death, then there should be an overabundance of downward lines on the right edge of the plots. That's not obvious, which is partly attributed to the substantial variation within individuals - activity measurements for a given copepod can differ quite a lot from day to day. Maybe the pattern emerges when we plot averages.

```{r}
# average values for dead
deads <- filter(bdr, dead==1, !is.na(speed_per_sec))%>%
  group_by(rec_time, days_surv, day_char)%>%
  summarize(speed_m = mean(speed_per_sec),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))

# average values for survivors
deads2 <- filter(bdr, dead==0)%>%
  group_by(rec_time, days_surv, day_char)%>%
  summarize(speed_m = mean(speed_per_sec),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))

# combine for plot
deads2 <- left_join(select(ungroup(deads), rec_time, day_char, days_surv), 
                    select(ungroup(deads2), -days_surv),
                    by = c('rec_time', 'day_char'))
deads2 <- filter(deads2, days_surv != 9)

ggplot(filter(deads, days_surv != 9),
       aes(x = as.numeric(day_char), y = speed_m)) +
  geom_point(color = 'red') +
  geom_errorbar(aes(ymin = lci, ymax = uci), 
                color = 'red', width = 0.5) +
  geom_line(color = 'red') +
  geom_point(data = deads2, aes(x = as.numeric(day_char), y = speed_m), color = 'gray') +
  geom_errorbar(data = deads2, aes(x = as.numeric(day_char), y = speed_m,
                                   ymin = lci, ymax = uci),
                color = 'gray', width = 0.5) +
  geom_line(data = deads2, aes(x = as.numeric(day_char), y = speed_m), color = 'gray') +
  facet_grid(days_surv~rec_time) +
  scale_x_continuous(breaks = c(5,7,9,11,13,15,17,19,21)) +
  labs(title = 'auto tracking - 95% CI',
       x = 'days post infection', y = 'average speed')
```

In the above plot, the average activity of copepods that died are red. I also plotted the values for copepods that did not die in gray for comparison. In some cases, copepod activity seems to drop as death approaches (for copepods that died on days 21, 20, 18), but not in others (days 16, 12, 10). Early in the experiment, when death is not imminent for most copepods, there is little difference between the dead and surviving copepods. Thus, my inclination is not to toss these data, as the differences seem minor, particularly because the 25 host-parasite genotype combinations cut the data relatively thin. Still, it also seems prudent to re-run the final analysis excluding these copepods to ensure that they do not bias the results.

# Treatment differences

Now let's compare the behavior of infected and uninfected copepods over the course of the experiment.

```{r}
bd_avg <- filter(bd, !is.na(speed_per_sec))%>%
  group_by(rec_time, trt, day_char)%>%
  summarize(speed_m = mean(speed_per_sec),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))

ggplot(bd_avg, aes(x = as.numeric(day_char), y = speed_m, color = trt)) +
  geom_point(position = position_dodge(width = 1)) +
  geom_errorbar(aes(ymin = lci, ymax = uci), 
                position = position_dodge(width = 1)) +
  geom_line(position = position_dodge(width = 1)) +
  scale_x_continuous(breaks = c(5,7,9,11,13,15,17)) +
  facet_grid(~rec_time) +
  labs(title = 'manual tracking', color = 'Treatment', x = 'days post infection', y = 'average speed')
```
```{r}
bdr_avg <- filter(bdr, !is.na(speed_per_sec))%>%
  group_by(rec_time, trt, day_char)%>%
  summarize(speed_m = mean(speed_per_sec),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))

ggplot(bdr_avg, aes(x = as.numeric(day_char), y = speed_m, color = trt)) +
  geom_point(position = position_dodge(width = 1)) +
  geom_errorbar(aes(ymin = lci, ymax = uci), 
                position = position_dodge(width = 1)) +
  geom_line(position = position_dodge(width = 1)) +
  facet_grid(~rec_time) +
  scale_x_continuous(breaks = c(5,7,9,11,13,15,17,19,21)) +
  labs(title = 'auto tracking', color = 'Treatment', x = 'days post infection', y = 'average speed')
```

Infected copepods are less active than uninfected copepods, regardless of days post infection. The worm becomes infective to its fish next host at around 11 days post infection. Other [studies](HAMMERSCHmidT) have found a conspicuous change in the behavior of infected copepods as worms develop from uninfective to infective larvae. Specifically they switch from reducing to increasing activity, relative to uninfecteds. However, this switch seems to be population-dependent, and it is not observed in the parasite population used in this experiment (see [here](HAFER)). In any case, the behavioral changes appear to affect copepods' predation susceptibility (see [here]()) and consequently parasite transmission and survival probabilities. Thus, the changes are likely to be fitness components under selection.

The response to the plate drop looks fairly consistent over time, though it is hard to gauge above. I plotted the change in activity in response to the drop below. The response seems rather consistent in the uninfected groups, while the infecteds seem to show less of a response to the drop towards the end of the experiment. However, this could also be because infecteds were slightly less active in general at the end of the experiment.

```{r}
ggplot(bdr_avg,
       aes(x = rec_time, y = speed_m, color = trt)) + 
  geom_line(aes(group = trt), position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = lci, ymax = uci), 
                position = position_dodge(width = 0.3)) +
  geom_point(position = position_dodge(width = 0.3)) +
  facet_wrap(~day_char, nrow = 3) +
  labs(title = 'Facetted by observation day', y = 'average speed', x = 'Drop', color = 'Treatment')
```

Our focus in this experiment is parasitic manipulation, i.e. how much infected copepods differ from uninfected copepods. But to look at genotype-by-genotype interactions we are restricted to using only infected copepods (uninfected copepods have no associated parasite genotype). So how should we define manipulation? I think there are two ways. First and simplest, we can take the raw behavior values for infected copepods, knowing that they differ from uninfecteds. Second and somewhat more complex, we can express infected copepod behavior relative to uninfected copepods, e.g. as a difference. In this first case, we assume that the absolute magnitude of behavior is most relevant for the parasite. In the second case, we consider manipulation in a relative sense in that some hosts at some times may be more or less susceptible to manipulation. We'll start with the simpler case - how do host and parasite genotypes affect infected copepod behavior?

# Manipulation in an absolute sense

Different copepod genotypes, which were inbred families in this experiment, exhibited different levels of activity. 

```{r}
bdr_avg <- filter(bdr, trt == 'infected')%>%
  group_by(cop_fam)%>%
  summarize(speed_m = mean(speed_per_sec, na.rm=T),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))
relab <- as.character(arrange(bdr_avg, speed_m)$cop_fam)
bdr_avg <- mutate(bdr_avg, cop_fam = factor(cop_fam, levels = relab))

ggplot(bdr_avg, aes(x = cop_fam, y = speed_m)) +
  geom_errorbar(aes(ymin = lci, ymax = uci)) +
  geom_point() +
  labs(title = 'auto tracking - confidence intervals underestimated', x = 'host genotype', y = 'average speed')
```

Similarly, different parasite genotypes, full-sib families in this case, also exhibit different average activities. The overlap between parasite genotypes appears smaller than that between hosts, which implies that parasite genotype has a larger effect than host genotype.

```{r}
bdr_avg <- filter(bdr, trt == 'infected')%>%
  group_by(parasite_fam)%>%
  summarize(speed_m = mean(speed_per_sec, na.rm=T),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))
relab <- as.character(arrange(bdr_avg, speed_m)$parasite_fam)
bdr_avg <- mutate(bdr_avg, parasite_fam = factor(parasite_fam, levels = relab))

ggplot(bdr_avg, aes(x = parasite_fam, y = speed_m)) +
  geom_errorbar(aes(ymin = lci, ymax = uci)) +
  geom_point() +
  labs(title = 'auto tracking - confidence intervals underestimated', x = 'parasite genotype', y = 'average speed')
```

Now, we can plot host and parasite genotype simultaneously, i.e. their interaction.

```{r}
bdr_avg <- filter(bdr, trt == 'infected')%>%
  group_by(cop_fam, parasite_fam)%>%
  summarize(speed_m = mean(speed_per_sec, na.rm=T),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)),
         parasite_fam = factor(parasite_fam, levels = relab))

ggplot(bdr_avg, aes(x = parasite_fam, y = speed_m, color = cop_fam)) +
  geom_line(aes(group = cop_fam), position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = lci, ymax = uci), 
                position = position_dodge(width = 0.3)) +
  geom_point(position = position_dodge(width = 0.3)) +
  labs(title = 'auto tracking - confidence intervals underestimated', x = 'parasite genotype', 
       y = 'average speed', color = 'host genotype')
```

There appear to be both main effects (e.g. parasite genotype B induces less activity than genotype C) and interactions (e.g. some host genotypes show quite different behavior depending on the parasite genotype). Judging from these initial plots, it looks like parasite genotype has a clear affect, host genotypes less so, and there are interactions. Modeling is needed to confirm this, though. In the last three plots, I noted that confidence intervals were underestimated, and this is because I pooled together the multiple observations from the same copepod across the experiment. Now let's explore whether the effects of host and parasite genotype are consistent over time.

```{r}
# cop fams
bdr_avg <- filter(bdr, trt == 'infected')%>%
  group_by(cop_fam, day_char)%>%
  summarize(speed_m = mean(speed_per_sec, na.rm=T))

ggplot(bdr_avg, aes(x = as.numeric(day_char), y = speed_m, color = cop_fam)) +
  geom_point(position = position_dodge(width = 0.75)) +
  geom_line(position = position_dodge(width = 0.75)) +
  scale_x_continuous(breaks = c(5,7,9,11,13,15,17,19,21)) +
  labs(title = 'auto tracking', color = 'Host genotype', x = 'days post infection', y = 'average speed')
```

The relative rank of copepod genotypes varies quite a lot. For example, genotype IV shows the lowest average activity on day 15 but the highest on day 21.

```{r}
# parasite fams
bdr_avg <- filter(bdr, trt == 'infected')%>%
  group_by(parasite_fam, day_char)%>%
  summarize(speed_m = mean(speed_per_sec, na.rm=T))

ggplot(bdr_avg, aes(x = as.numeric(day_char), y = speed_m, color = parasite_fam)) +
  geom_point(position = position_dodge(width = 1)) +
  geom_line(position = position_dodge(width = 1)) +
  scale_x_continuous(breaks = c(5,7,9,11,13,15,17,19,21)) +
  labs(title = 'auto tracking', color = 'Parasite genotype', x = 'days post infection', y = 'average speed')
```

The differences among parasite genotypes appears more consistent, which is consistent with the idea that parasite genotype affects behavior more than host genotype. However, the relative rank of different groups still change over time, indicating temporal effects that need to be considered in the model.






### Models

Here is my model-building approach. First, establish a base model with the variables of non-primary interest, including day post-infection, whether activity is before or after the drop, individual copepod, and block.

```{r}
# start with linear mixed models, before experimenting with Tweedie or Gamma
library(lme4)
# testing base models
lm0 <- lmer(speed_per_sec ~ rec_time + day_char + (1|cop_name) + (1|box), # no interactions
            data = filter(bdr, trt == 'infected'))
lm0.1 <- lmer(speed_per_sec ~ rec_time * day_char + (1|cop_name) + (1|box), # let reaction to drop depend on dpi
            data = filter(bdr, trt == 'infected'))
lm0.2 <- lmer(speed_per_sec ~ rec_time * day_char + 
                (1|cop_name) + (1|box) + (rec_time|cop_name), # let reaction to drop vary by copepod - rand slopes
            data = filter(bdr, trt == 'infected'))
anova(lm0, lm0.1, lm0.2) # adding time-dependent drop response improves model, but copepod-dependent drop response does not
```

Then, we add copepod and worm genotypes to the model. Has it improved?

```{r}
lm1.0 <- lmer(speed_per_sec ~ rec_time * day_char + cop_fam +
                (1|cop_name) + (1|box), # add cop fam
            data = filter(bdr, trt == 'infected'))
lm1.1 <- lmer(speed_per_sec ~ rec_time * day_char + parasite_fam +
                (1|cop_name) + (1|box), # add parasite fam
            data = filter(bdr, trt == 'infected'))
lm1 <- lmer(speed_per_sec ~ rec_time * day_char + cop_fam + parasite_fam +
                (1|cop_name) + (1|box), # add cop and parasite fam
            data = filter(bdr, trt == 'infected'))
anova(lm0.1, lm1.0, lm1)
anova(lm0.1, lm1.1, lm1) # adding both is an improvement
```

Next, we add the interaction between host and parasite. It is significant.

```{r}
lm2 <- lmer(speed_per_sec ~ rec_time * day_char + cop_fam * parasite_fam +
                (1|cop_name) + (1|box), # add GxG interaction
            data = filter(bdr, trt == 'infected'))
anova(lm1, lm2) # interaction is sig
```

Next, I allow host-parasite interactions to vary across days of the experiment. It is significant, suggesting substantial temporal variation in genotype effects.

```{r}
lm3 <- lmer(speed_per_sec ~ rec_time * day_char + cop_fam * parasite_fam * day_char +
                (1|cop_name) + (1|box), # let GxG combos behave differently over time
            data = filter(bdr, trt == 'infected'))
anova(lm2, lm3)
```

Finally, let host-parasite interactions vary in their response to the drop across days of the experiment. Not significant.

```{r}
lm4 <- lmer(speed_per_sec ~ rec_time * day_char + cop_fam * parasite_fam * day_char * rec_time +
                (1|cop_name) + (1|box), # let GxG combos react to drop differently on each dpi
            data = filter(bdr, trt == 'infected'))
anova(lm3, lm4)
```

So, there are both genotype main effects and interactions on behavior. The response to the drop does not seem genotype-dependent. There is also substantial variation from one day to the next. Perhaps some of the temporal variation decreases when we consider manipulation in a relative sense. 

# Manipulation in a relative sense 

Combine uninfecteds, because no diff.

```{r}
# behav over time in uninfected copepods
bdr_avg <- filter(bdr, infected == 0)%>%
  group_by(cop_fam, day_char)%>%
  summarize(speed_m = mean(speed_per_sec, na.rm=T),
            speed_sd = sd(speed_per_sec),
            n = n())%>%
  mutate(uci = speed_m + 2 * (speed_sd/sqrt(n)),
         lci = speed_m - 2 * (speed_sd/sqrt(n)))

ggplot(bdr_avg, aes(x = as.numeric(day_char), y = speed_m, color = cop_fam)) + 
  geom_line(aes(group = cop_fam), position = position_dodge(width = 0.3)) +
  #geom_errorbar(aes(ymin = lci, ymax = uci), 
   #             position = position_dodge(width = 0.3)) +
  geom_point(position = position_dodge(width = 0.3)) +
  labs(title = 'uninfected copepods', x ="days post infection",
       y = 'average speed', color = 'host genotype')
```

Bigger effect size of cop fam than in the infecteds only.

```{r}
lm0.2 <- lmer(speed_per_sec ~ rec_time * day_char + (1|cop_name) + (1|box), # let reaction to drop depend on dpi
            data = filter(bdr, infected == 0))
lm1.2 <- lmer(speed_per_sec ~ rec_time * day_char + cop_fam +
                (1|cop_name) + (1|box), # let reaction to drop depend on dpi
            data = filter(bdr, infected == 0))
anova(lm0.2, lm1.2) # seems to be 
```

## Relative to uninfecteds

## Relative to uninfecteds in specific family

# Variance partitioning (or comparing effect sizes)

# Correlated traits
GxG in infection, development, growth

Rerun excluding dead
Rerun basics with manual data