---
title: "Analyses in ms"
author: "Dan Benesh"
date: "6/24/2019"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)
theme_set(new = theme_bw())
```

# Experiment Background

Many parasites manipulate the behavior of their hosts in ways that seem beneficial. Little is known about genetic variation in this phenotype. Are parasites genetically variable in how much they manipulate host behavior? Do different host genotypes resist manipulation better than others? I infected 5 host strains with 5 parasite strains to quantify the variability in host manipulation due to parasite genes, host genes, and their interactions. The host-parasite system I used was the tapeworm *Schistocephalus solidus* in its copepod first intermediate host.

```{r importdata}
# import data - averaged within recordings
bd <- read.csv(file = "../data/behav_manutracked_reduced_dataset.csv", header = TRUE) # manual data
bdr <- read.csv(file = "../data/behav_autotracked_reduced_dataset.csv", header = TRUE) # auto-tracked data
infd <- read.csv(file = "../data/GxG_inf.csv", header = TRUE, fileEncoding = "UTF-8-BOM") #excel attached BOM to csv
```

```{r}
# take average worm size
infd <- mutate(infd, proc_size = (proc_size1 + proc_size2)/2)
```
```{r}
# add useful variables to manual data
bd <- mutate(bd, speed_per_sec = tot_dist/frames_tracked * 0.5,
       cop_name = substring(fname, 1, 5),
       day_char = substring(fname, 7, 8),
       plate_no = substring(fname, 1, 2),
       rec_time = factor(group, levels = c('before the drop', 'after the drop')))%>%
  select(-group)

# combine with inf data
bd <- left_join(bd, select(infd, -proc_size1, -proc_size2, -exposed, -tank, -cerc_multinom, -remarks),
                by = 'cop_name')
# better labels to trts
bd <- mutate(bd, trt = factor(trt, levels = c('unexposed', 'uninfected', 'infected')),
       trt = factor(trt, labels = c('unexposed, control', 'exposed, uninfected', 'infected')),
       parasite_fam = factor(parasite_fam, labels = c('A', 'B', 'C', 'D', 'E')),
       box = factor(box),
       devo_period = if_else(day_char %in% c("05", "07", "09"), "uninfective", 
                             if_else(day_char %in% c("11", "13", "15"), "partially infective", "infective")))%>%
  mutate(devo_period = factor(devo_period, levels = c("uninfective", "partially infective", "infective")))
```
```{r}
# add useful variables to auto data
bdr <- mutate(bdr, speed_per_sec = tot_dist/frames_tracked * 2,
       cop_name = substring(fname, 1, 5),
       day_char = substring(fname, 7, 8),
       plate_no = substring(fname, 1, 2),
       rec_time = factor(group, levels = c('before the drop', 'after the drop')))%>%
  select(-group)


# combine with inf data
bdr <- left_join(bdr, select(infd, -proc_size1, -proc_size2, -exposed, -tank, -cerc_multinom, -remarks),
                by = 'cop_name')
# better labels to trts
bdr <- mutate(bdr, trt = factor(trt, levels = c('unexposed', 'uninfected', 'infected')),
              trt = factor(trt, labels = c('unexposed, control', 'exposed, uninfected', 'infected')),
              parasite_fam = factor(parasite_fam, labels = c('A', 'B', 'C', 'D', 'E')),
              box = factor(box),
              devo_period = if_else(day_char %in% c("05", "07", "09"), "uninfective", 
                                    if_else(day_char %in% c("11", "13", "15"), "partially infective", "infective")))%>%
  mutate(devo_period = factor(devo_period, levels = c("uninfective", "partially infective", "infective")))
```
```{r}
bdr <- mutate(bdr, infection = if_else(infected == 0, "uninfected", "infected"))%>% # make a nice variable for infection
  mutate(infection = factor(infection, levels = c("uninfected", "infected")))
```
```{r}
bdr <- mutate(bdr, parasite_fam2 = if_else(infected == 0, "uninfected", as.character(parasite_fam))) # re-formulate parasite family factor so all uninfecteds have same value
```
```{r}
bdr <- mutate(bdr, proc2 = proc_size/1000) # rescale to avoid numeric problems
```

```{r}
library(cplm)
```

# Combining controls and exposed-but-uninfected copepods

```{r}
cpm_uninf1 <- cpglmm(speed_per_sec ~ box + rec_time * day_char + (1|cop_name), 
                     data = filter(bdr, trt != 'infected', !is.na(trt)))
cpm_uninf2 <- update(cpm_uninf1, . ~ . + trt)

anova(cpm_uninf1, cpm_uninf2)
```

# Role of infection - Table 1


We'll fit these models using the same approach as with the linear mixed models. First initialize them, and find the best 'base' model.

```{r}
# initialize models
cpm_int <- cpglmm(speed_per_sec ~ 1 + (1|cop_name), 
            data = bdr)
cpm0 <- update(cpm_int, . ~ . + box + rec_time + day_char) # base model
cpm01 <- update(cpm0, . ~ . + rec_time * day_char) # time-dep drop resp

anova(cpm_int, cpm0, cpm01)
```

Then, we assess the role of infection.

```{r}
cpm1 <- update(cpm01, . ~ . + infection) # infection
cpm11 <- update(cpm1, . ~ . + infection * rec_time) # drop resp x infection
cpm12 <- update(cpm11, . ~ . + infection * day_char) # day x inf
cpm13 <- update(cpm12, . ~ . + infection * day_char * rec_time) # drop x day x inf

anova(cpm01, cpm1, cpm11, cpm12, cpm13)
```

Next, we examine copepod genotype effects.

```{r}
cpm2 <- update(cpm12, . ~ . + cop_fam) # cop genotype
cpm21 <- update(cpm2, . ~ . + cop_fam * infection) # cop geno x infection
cpm22 <- update(cpm21, . ~ . + cop_fam * infection * rec_time) # cop geno drop resp

anova(cpm12, cpm2, cpm21, cpm22 )
```

```{r}
r2_cglmm <- function(null_model, model) {
  # take in compound poisson mixed model, return marginal and conditional R2, ala Nakagawa et al. 2017
  # marginal r2 is just fixed effects
  # condition r2 is fixed and rand effects combined
  
  # model call
  call <- as.character(model@call)[2]
  
  # parameter estimates and df
  fixed_param <- fixef(model)
  df <- length(fixed_param) - 1
  
  # variance due to fixed effects
  pred <- as.vector(model.matrix(model) %*% fixed_param) # predicteds on basis of just fixed effects
  varF <- var(pred)
  
  
  # getting the observation-level variance of null model...
  phiN <- null_model@phi # the dispersion parameter
  pN <- null_model@p # the index parameter
  mu <- exp(fixef(null_model) + 0.5 * (VarCorr(null_model)$cop_name[1]))
  # ...use to estimate dist-specific resid variance from full
  phiF <- model@phi # the dispersion parameter 
  pF <- model@p # the index parameter 
  VarOdF <- phiF * mu^(pF - 2) # the delta method 
  VarOlF <- log(1 + (phiF * mu^(pF - 2))) # log-normal approximation 
  
  # variance due to rand effects
  var_cop <- VarCorr(model)$cop_name[1]
  
  
  # marginal r2
  mr2 <- varF/(varF + var_cop + VarOdF)
  
  # conditional r2
  cr2 <- (varF + var_cop)/(varF + var_cop + VarOdF)
  
  # output
  out_frame <- data_frame(call = call, df = df, marg_r2 = round(mr2, 3), cond_r2 = round(cr2,3))
  return(out_frame)
}
```

```{r}
mod_list <- list(cpm_int, cpm0, cpm01, cpm1, cpm11, cpm12, cpm13, cpm2, cpm21, cpm22)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_cglmm(cpm_int, model)
  } else {
    r2_table <- rbind(r2_table, r2_cglmm(cpm_int, model))
  }
  i <- i + 1
}

r2_table <- mutate(r2_table, cop_ind_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("intercept", "base", "base2", "infection", "infection x drop", "infection x day", "infection x day x drop",
                   "cop genotype", "cop x infection", "cop genotype x drop")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, cop_ind_var_explained)
r2_table
```


Means for figures

```{r}
cpm_figs1 <- update(cpm_int, . ~ . + day_char:trt - 1) # means in Fig S1
```
```{r}
cpm_figs2 <- update(cpm_int, . ~ . + day_char:trt:rec_time - 1) # means in Fig S2
```
```{r}
cpm_fig1 <- update(cpm_int, . ~ . + cop_fam:infection - 1) # means in Fig 1
```


# Variation within infected copepods - Table 1

```{r}
cpmi_int <- cpglmm(speed_per_sec ~ 1 + (1|cop_name), 
            data = filter(bdr, infection == 'infected'))

cpmi0 <- update(cpmi_int, . ~ . + box + rec_time + day_char) # base model
cpmi01 <- update(cpmi0, . ~ . + rec_time * day_char) # time-dep drop resp

anova(cpmi_int, cpmi0, cpmi01 )
```

```{r}
cpmi2 <- update(cpmi01, . ~ . + cop_fam) # cop genotype
cpmi21 <- update(cpmi2, . ~ . + cop_fam * rec_time) # cop geno drop resp

anova(cpmi01, cpmi2, cpmi21 )
```

```{r}
cpmi3 <- update(cpmi2, . ~ . + parasite_fam) # worm genotype
cpmi31 <- update(cpmi3, . ~ . + parasite_fam * rec_time) # worm geno drop resp

anova(cpmi2, cpmi3, cpmi31 )
```

```{r}
cpmi4 <- update(cpmi3, . ~ . + cop_fam * parasite_fam) # G x G
cpmi5 <- update(cpmi4, . ~ . + cop_fam * parasite_fam * day_char) # G x G x time

anova(cpmi3, cpmi4, cpmi5)
```

```{r}
mod_list <- list(cpmi_int, cpmi0, cpmi01, cpmi2, cpmi21, cpmi3, cpmi31, cpmi4, cpmi5)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_cglmm(cpmi_int, model)
  } else {
    r2_table <- rbind(r2_table, r2_cglmm(cpmi_int, model))
  }
  i <- i + 1
}

r2_table <- mutate(r2_table, cop_ind_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("intercept", "base", "base2", 
                   "cop genotype", "cop genotype x drop",
                   "worm genotype", "worm geno x drop", 
                   "G x G", "G x G x time")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, cop_ind_var_explained)
r2_table
```


Means for figs

```{r}
cpm_fig2a <- update(cpmi_int, . ~ . + cop_fam - 1) # means in Fig 2A
```
```{r}
cpm_fig2b <- update(cpmi_int, . ~ . + parasite_fam - 1) # means in Fig 2B
```
```{r}
cpm_fig2b2 <- update(cpmi_int, . ~ . + parasite_fam + cop_fam) # means in combined version Fig 2B
```
```{r}
cpm_fig2c <- update(cpmi_int, . ~ . + cop_fam:parasite_fam - 1) # means in Fig 2C
```


# Variation within uninfected copepods - cop family explains less variation in uninf than in inf copepods?

```{r}
cpmu_int <- cpglmm(speed_per_sec ~ 1 + (1|cop_name), 
            data = filter(bdr, infection != 'infected'))

cpmu0 <- update(cpmu_int, . ~ . + box + rec_time + day_char) # base model
cpmu01 <- update(cpmu0, . ~ . + rec_time * day_char) # time-dep drop resp

anova(cpmu_int, cpmu0, cpmu01 )
```

```{r}
cpmu2 <- update(cpmu01, . ~ . + cop_fam) # cop genotype
cpmu21 <- update(cpmu2, . ~ . + cop_fam * rec_time) # cop geno drop resp

anova(cpmu01, cpmu2, cpmu21 )
```

```{r}
mod_list <- list(cpmu_int, cpmu0, cpmu01, cpmu2, cpmu21)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_cglmm(cpmu_int, model)
  } else {
    r2_table <- rbind(r2_table, r2_cglmm(cpmu_int, model))
  }
  i <- i + 1
}

r2_table <- mutate(r2_table, cop_ind_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("intercept", "base", "base2", 
                   "cop genotype", "cop genotype x drop")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, cop_ind_var_explained)
r2_table
```


# Same models, dead excluded

## Infected + uninfected cops

```{r}
cpmd_int <- cpglmm(speed_per_sec ~ 1 + (1|cop_name), 
            data = filter(bdr, dead == 0))
cpmd0 <- update(cpmd_int, . ~ . + box + rec_time + day_char) # base model
cpmd01 <- update(cpmd0, . ~ . + rec_time * day_char) # time-dep drop resp

anova(cpmd_int, cpmd0, cpmd01)
```

Then, we assess the role of infection.

```{r}
cpmd1 <- update(cpmd01, . ~ . + infection) # infection
cpmd11 <- update(cpmd1, . ~ . + infection * rec_time) # drop resp x infection
cpmd12 <- update(cpmd11, . ~ . + infection * day_char) # day x inf
cpmd13 <- update(cpmd12, . ~ . + infection * day_char * rec_time) # day x inf
anova(cpmd01, cpmd1, cpmd11, cpmd12, cpmd13)
```

Next, we examine copepod genotype effects.

```{r}
cpmd2 <- update(cpmd12, . ~ . + cop_fam) # cop genotype
cpmd21 <- update(cpmd2, . ~ . + cop_fam * infection) # cop geno x infection
cpmd22 <- update(cpmd21, . ~ . + cop_fam * infection * rec_time) # cop geno drop resp

anova(cpmd12, cpmd2, cpmd21, cpmd22 )
```

```{r}
mod_list <- list(cpmd_int, cpmd0, cpmd01, cpmd1, cpmd11, cpmd12, cpmd13, cpmd2, cpmd21, cpmd22)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_cglmm(cpmd_int, model)
  } else {
    r2_table <- rbind(r2_table, r2_cglmm(cpmd_int, model))
  }
  i <- i + 1
}

r2_table <- mutate(r2_table, cop_ind_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("intercept", "base", "base2", 
                   "infection", "infection x drop", "infection x day", "infection x drop x day", 
                   "cop genotype", "cop x infection", "cop genotype x drop")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, cop_ind_var_explained)
r2_table
```

## Just infecteds

```{r}
cpmid_int <- cpglmm(speed_per_sec ~ 1 + (1|cop_name), 
            data = filter(bdr, dead == 0, infection == 'infected'))

cpmid0 <- update(cpmid_int, . ~ . + box + rec_time + day_char) # base model
cpmid01 <- update(cpmid0, . ~ . + rec_time * day_char) # time-dep drop resp

anova(cpmid_int, cpmid0, cpmid01 )
```

```{r}
cpmid2 <- update(cpmid01, . ~ . + cop_fam) # cop genotype
cpmid21 <- update(cpmid2, . ~ . + cop_fam * rec_time) # cop geno drop resp

anova(cpmid01, cpmid2, cpmid21 )
```

```{r}
cpmid3 <- update(cpmid2, . ~ . + parasite_fam) # worm genotype
cpmid31 <- update(cpmid3, . ~ . + parasite_fam * rec_time) # worm geno drop resp

anova(cpmid2, cpmid3, cpmid31 )
```

```{r}
cpmid4 <- update(cpmid3, . ~ . + cop_fam * parasite_fam) # G x G
cpmid5 <- update(cpmid4, . ~ . + cop_fam * parasite_fam * day_char) # G x G x time

anova(cpmid3, cpmid4, cpmid5)
```

```{r}
mod_list <- list(cpmid_int, cpmid0, cpmid01, cpmid2, cpmid21, cpmid3, cpmid31, cpmid4, cpmid5)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_cglmm(cpmi_int, model)
  } else {
    r2_table <- rbind(r2_table, r2_cglmm(cpmi_int, model))
  }
  i <- i + 1
}

r2_table <- mutate(r2_table, cop_ind_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("intercept", "base", "base2", 
                   "cop genotype", "cop genotype x drop",
                   "worm genotype", "worm geno x drop", 
                   "G x G", "G x G x time")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, cop_ind_var_explained)
r2_table
```

# Manual data

## Infected + uninfected cops

```{r}
bd <- mutate(bd, infection = if_else(infected == 0, "uninfected", "infected"), 
             parasite_fam2 = if_else(infected == 0, "uninfected", as.character(parasite_fam)))%>% # re-formulate parasite family factor so all uninfecteds have same value
  mutate(infection = factor(infection, levels = c("uninfected", "infected")))

cpmm_int <- cpglmm(speed_per_sec ~ 1 + (1|cop_name), # int only model
                data = bd)
cpmm0 <- update(cpmm_int, . ~ . + box + rec_time + day_char) # base model
cpmm01 <- update(cpmm0, . ~ . + rec_time * day_char) # base model2

anova(cpmm_int, cpmm0, cpmm01) 
```
```{r}
# add infection
cpmm1 <- update(cpmm01, . ~ . + infection)
cpmm11 <- update(cpmm1, . ~ . + infection * rec_time) 
cpmm12 <- update(cpmm11, . ~ . + infection * day_char) 
cpmm13 <- update(cpmm12, . ~ . + infection * day_char * rec_time) 

anova(cpmm01, cpmm1, cpmm11, cpmm12, cpmm13)
```

```{r}
# add cop fam
cpmm2 <- update(cpmm12, . ~ . + cop_fam) # copepod fam
cpmm21 <- update(cpmm2, . ~ . + infection * cop_fam) # copepod fam differ by infection
cpmm22 <- update(cpmm21, . ~ . + infection * cop_fam * rec_time) # copepod fam differ by infection

anova(cpmm12, cpmm2, cpmm21, cpmm22)
```

```{r}
mod_list <- list(cpmm_int, cpmm0, cpmm01, cpmm1, cpmm11, cpmm12, cpmm13, cpmm2, cpmm21, cpmm22)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_cglmm(cpmd_int, model)
  } else {
    r2_table <- rbind(r2_table, r2_cglmm(cpmd_int, model))
  }
  i <- i + 1
}

r2_table <- mutate(r2_table, cop_ind_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("intercept", "base", "base2", 
                   "infection", "infection x drop", "infection x day", "infection x drop x day", 
                   "cop genotype", "cop x infection", "cop genotype x drop")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, cop_ind_var_explained)
r2_table
```

## Just infecteds

```{r}
cpmim_int <- cpglmm(speed_per_sec ~ 1 + (1|cop_name), # int only model
                data = filter(bd, infection == "infected"))
cpmim0 <- update(cpmim_int, . ~ . + box + rec_time + day_char) # base model
cpmim01 <- update(cpmim0, . ~ . + rec_time * day_char) # base model2

anova(cpmim_int, cpmim0, cpmim01)
```

```{r}
cpmim2 <- update(cpmim01, . ~ . + cop_fam) # copepod fam
cpmim21 <- update(cpmim2, . ~ . + cop_fam * rec_time) # copepod fam x drop

anova(cpmim01, cpmim2, cpmim21)
```

```{r}
cpmim3 <- update(cpmim2, . ~ . + parasite_fam) # worm fam
cpmim31 <- update(cpmim3, . ~ . + parasite_fam * rec_time) # worm fam x drop

anova(cpmim2, cpmim3, cpmim31)
```


```{r}
cpmim4 <- update(cpmim3, . ~ . + cop_fam * parasite_fam) # GxG
cpmim5 <- update(cpmim4, . ~ . + cop_fam * parasite_fam * day_char) # time-dependent

anova(cpmim3, cpmim4, cpmim5)
```

```{r}
mod_list <- list(cpmim_int, cpmim0, cpmim01, cpmim2, cpmim21, cpmim3, cpmim31, cpmim4, cpmim5)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_cglmm(cpmi_int, model)
  } else {
    r2_table <- rbind(r2_table, r2_cglmm(cpmi_int, model))
  }
  i <- i + 1
}

r2_table <- mutate(r2_table, cop_ind_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("intercept", "base", "base2", 
                   "cop genotype", "cop genotype x drop",
                   "worm genotype", "worm geno x drop", 
                   "G x G", "G x G x time")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, cop_ind_var_explained)
r2_table
```