---
title: "Genotype x genotype interactions in host manipulation"
author: "Dan Benesh"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo = FALSE)
options(stringsAsFactors = FALSE)
```

# Background
Many parasites manipulate the behavior of their hosts in ways that seem beneficial. Little is known about genetic variation in this phenotype. Are parasites genetically variable in how much they manipulate host behavior? Do different host genotypes resist manipulation better than others? I infected 5 host strains with 5 parasite strains to quantify the variability in host manipulation due to parasite genes, host genes, and their interactions. The host-parasite system I used was the tapeworm *Schistocephalus solidus* in its copepod first intermediate host.

```{r importdata}
# import data
bd <- read.csv(file = "../data/behav_combined_after_qc2.csv", header = TRUE) # manual data
bdr <- read.csv(file = "../data/behav_combined_auto_tracked_corrected.csv", header = TRUE) # auto-tracked data
infd <- read.csv(file = "../data/GxG_inf.csv", header = TRUE, fileEncoding = "UTF-8-BOM") #excel attached BOM to csv
```

Copepod movement was video recorded for two minutes. After one minute, the well plate containing the copepods was dropped in a standardized way to simulate a predator attack and 'frighten' the copepods. Each copepod was recorded on several days throughout the course of parasite development (from uninfective to infective larva). 

Recordings were both manually- (clicking on copepod position frame-by-frame) and automatically-processed (with a tracking program). The output from the two approaches was quite comparable (see an extensive comparison [here](GITHUB LINK!!)). The auto-tracked data has some advantages, though. First, I identified mistakes in the manually-processed data, like incorrect number of frames, etc. (see [here](GITHUB LINK)); this motivated the development of the auto-tracker. Second, the auto-tracker allowed additional videos and copepods to be processed that were not included in the manual data for lack of time. I prefer the auto-tracked data as it is more standardized, but here I make sure that both datasets yield concordant results.

In this notebook, I first present descriptive statistics from the experiment. Then I look at ways of defining the response variable. I conduct exploratory analyses of the response variable. Finally, I fit models to test the hypothesis of genotype-by-genotype interactions.

```{r}
# Select and rename columns
bd <- select(bd, fname, cop_name, day, 
             slice = Slice.n., x = X, y = Y,
             dist = Distance, pixel = Pixel.Value, 
             ok_col_names, ok_col_num, ok_row_num)%>%
  arrange(day, cop_name, slice)

# for first observation of every video, no 'distance moved' can be calculated; replace with NA
bd$dist[bd$slice == 1 & bd$dist == -1] <- NA

# create new variables
bd <- mutate(bd, time_sec = slice * 2 - 4)
```

# Descriptive Statistics

Let's look at some basic stats for the infection. How many copepods were in the experiment?

```{r}
# correct problem from import with `cop_name` column
if(grepl(pattern = "?..", names(infd)[1])) { # if first column name has weird '?..' after import...
  names(infd)[1] <- 'cop_name' # change it
} 

# create new trt var with unexposed and uninfecteds pooled
infd$trt2 <- infd$trt
infd$trt2[which(infd$infected == 0)] <- 'uninfected'
```
```{r}
sum(!is.na(infd$exposed)) # if included, cops had a value for 'exposed' variable
```

Most of these were exposed to infection. 

```{r}
sum(infd$exposed == 1, na.rm=T)
```

The rest were unexposed controls (n = `r sum( infd$exposed == 0, na.rm=T)`).

Whether copepods were infected or not was checked after a week. How many survived until being checked for infection?

```{r}
sum( !is.na(infd$infected) )
```

And how many were infected?

```{r}
sum(infd$infected == 1, na.rm=T)
```

So the infection rate was `r round(sum(infd$infected == 1, na.rm=T) / sum( infd$exposed == 1, na.rm=T), 3) * 100`%.

We can also do some quality control checks on our infection data. For example, I measured some parasite traits like cercomere presence (an indication of developmental rate) and worm size. So we can check to make sure these variables are only recorded from infected copepods, starting with cercomere presence. 

It is, though 5 infected copepods are missing cercomere data.
```{r}
table(!is.na(infd$cerc_d9), infd$trt)
```

Similarly, worm size was only recorded from infected copepods, though in 99 it was not measured. Most of those without size measurement died before the end of the experiment 21 days post exposure.

```{r}
table(!is.na(infd$proc_size1), infd$trt)
```

Worm size was measured twice. The two measures are highly correlated, with a pearson correlation coefficient of `r round( with(infd, cor(proc_size1, proc_size2, use = "pairwise")), 2)`, though there was some variation in the measurements.

```{r message=FALSE, warning=FALSE}
qplot(data = infd, x = proc_size1, y = proc_size2) + theme_bw()
```

Presumably, this is normal measurement error due to e.g. photo angles, moving worms, etc. The average of the two body size measurements is taken as the best measure of larval size in future analyses.

```{r}
infd <- mutate(infd, proca = (proc_size1 + proc_size2)/2) # average of 2 proc measurements
```

Moving on to the behavior data, how many copepods had their behavior recorded manually?

```{r}
bd_cops <- unique(bd$cop_name)
length(bd_cops)
```

How many were recorded automatically?

```{r}
bdr <- mutate(bdr, cop_name = substr(fname, 1, 5))
length(unique(bdr$cop_name))
```

There are 5 cop families and 5 parasite families. How many infected cops are in each group?

```{r}
# reduce infection table to just cops with behav measurements
infd_bd <- filter(infd, cop_name %in% bd_cops)

# make table
t <- with(filter(infd_bd, trt2 == 'infected'),
     table(cop_fam, parasite_fam))
t
```

It ranges from `r min(t)` to `r max(t)` per combination, with an average of `r mean(t)`.

What about uninfecteds? Let's check the numbers of unexposed controls and exposed-but-uninfected copepods in each copepod family. Generally about 30 per family.

```{r}
# make table
t <- with(filter(infd_bd, infected == 0),
     table(cop_fam, trt))
t
```

# Define Response Variables

Behavior was recorded over two minutes, with the plate being dropped in the middle of the recording. So *a priori* we can divide the recording into three groups: before the drop, during the drop, and after the drop. The simplest behavioral measurement is the distance moved between frames. Let's look at movement distance changes throughout a recording.

```{r}
bdr_avg <- group_by(bdr, sec)%>%
  summarize(mean_dist = mean(distance, na.rm = T),
            mean_dp = mean(dot_product, na.rm = T))%>%
  mutate(group = if_else(sec < 58.5, 'before the drop', 
                         if_else(sec > 61.5, 'after the drop', 'drop')))

ggplot(bdr_avg, aes(x = sec, y = mean_dist)) + 
  geom_path() + geom_point(aes(color=group)) +
  scale_x_continuous(limits = c(0,120), breaks = seq(0, 120, by = 10)) +
  labs(title = 'auto tracker, 8 frames/s')
```
```{r warning=FALSE}
bd_avg <- group_by(bd, time_sec)%>%
  summarize(mean_dist = mean(dist, na.rm=T))%>%
  mutate(group = if_else(time_sec < 59, 'before the drop', 
                         if_else(time_sec > 62, 'after the drop', 'drop')))
  
ggplot(bd_avg, aes(x = time_sec, y = mean_dist)) + 
  geom_path() + geom_point(aes(color=group)) +
  scale_x_continuous(limits = c(0,120), breaks = seq(0, 120, by = 10)) +
  labs(title = 'manual tracker, 1 frame/2 s')
```

The most conspicuous pattern is the spike at 60 second during the drop. The copepod moves so fast during this drop that it cannot be tracked. The slight change in plate position through the drop also causes problems with background subtraction in the auto-tracking program, so these values cannot be considered accurate and need to be removed before analyses. But how much should we remove on either side of the drop? 

Let's zoom in on the drop. It seems like taking the 1.5 seconds on either side of the drop seems to reasonably exclude most of the peak.

```{r}
ggplot(bdr_avg, aes(x = sec, y = mean_dist)) + 
  geom_path() + geom_point(aes(color=group)) +
  labs(title = 'auto tracker, 8 frames/s') + 
  scale_y_log10() + coord_cartesian(xlim = c(50, 70))
```

When we exclude the unreliable data around the 'drop', we see clearer patterns. 

```{r}
ggplot(filter(bdr_avg, group != 'drop'), 
       aes(x = sec, y = mean_dist, color = group)) + 
  geom_path() + 
  scale_x_continuous(limits = c(0,120), breaks = seq(0, 120, by = 10)) +
  labs(title = 'auto tracker, 8 frames/s')
```
```{r}
ggplot(filter(bd_avg, group != 'drop'), 
       aes(x = time_sec, y = mean_dist, color = group)) + 
  geom_path() + 
  scale_x_continuous(limits = c(0,120), breaks = seq(0, 120, by = 10)) +
  labs(title = 'manual tracking')
```

In both cases, there is a clear decrease in activity directly after the drop. However, the copepods seem to recover quickly and by the end of the second minute their activity levels are back up to those before the drop. Also, in the auto-tracked data, there is still a spike shortly before the drop. This might be problems with the tracker, given that it was not observed in the manual data.

Direction moved

Calculate angles of movements.

```{r}
# calculate angle of movement for auto-tracked data
bdr <- mutate(bdr, x0 = lag(x), x2 = lead(x), y0 = lag(y), y2 = lead(y))%>%
    mutate(dist1 = sqrt((x0 - x)^2 + (y0 - y)^2) )%>%
    mutate(dist2 = sqrt((x - x2)^2 + (y - y2)^2) )%>%
    mutate(angle = acos(dot_product/(dist1 * dist2)) * 180/pi)%>%
    select(fname, frame, sec, x, y, blobs, blob_size, 
           distance, dot_product, angle)

# remove the angle calc for first observation in each recording
x <- group_by(bdr, fname)%>%
  summarize(frame_min = min(frame))%>%
  mutate(mv = paste(fname, frame_min))

bdr <- mutate(bdr, mv = paste(fname, frame))
bdr$angle[which(bdr$mv %in% x$mv)] <- NA
bdr <- select(bdr, -mv)
```

```{r}
# calculate angle of movement for manually-tracked data
bd <- mutate(bd, x0 = lag(x), x2 = lead(x), y0 = lag(y), y2 = lead(y))%>%
    mutate(dist1 = sqrt((x0 - x)^2 + (y0 - y)^2) )%>%
    mutate(dist2 = sqrt((x - x2)^2 + (y - y2)^2) )%>%
    mutate(dot_product = ((x2 - x) * (x - x0) + (y2 - y) * (y - y0)))%>%
    mutate(angle = acos(dot_product/(dist1 * dist2)) * 180/pi)%>%
    select(fname, day, slice, time_sec, x, y, 
           dist, dot_product, angle)

head(bd)
# remove the angle, dot_product calc for first observation in each recording
x <- group_by(bd, fname)%>%
  summarize(frame_min = min(slice))%>%
  mutate(mv = paste(fname, frame_min))
bd <- mutate(bd, mv = paste(fname, slice))
bd$angle[which(bd$mv %in% x$mv)] <- NA
bd$dot_product[which(bd$mv %in% x$mv)] <- NA
bd <- select(bd, -mv)
```

Plot over time. Typical angle is super large, presumably due to small oscillations.

```{r}
bdr_avg <- group_by(bdr, sec)%>%
  summarize(median_dir = median(angle, na.rm = T),
            var_dir = var(angle, na.rm = T))%>%
  mutate(group = if_else(sec < 58.5, 'before the drop', 
                         if_else(sec > 61.5, 'after the drop', 'drop')))

ggplot(bdr_avg, aes(x = sec, y = median_dir)) + 
  geom_path() + geom_point(aes(color=group)) +
  scale_x_continuous(limits = c(0,120), breaks = seq(0, 120, by = 10)) +
  labs(title = 'Angle - auto tracker, 8 frames/s')
```

In manually-tracked lower angle, corresponding more to typical movements along the edge of the well. Increase after drop suggests change of direction.
```{r}
bd_avg <- group_by(bd, time_sec)%>%
  summarize(median_dir = median(angle, na.rm=T),
            var_dir = var(angle, na.rm=T))%>%
  mutate(group = if_else(time_sec < 59, 'before the drop', 
                         if_else(time_sec > 62, 'after the drop', 'drop')))

ggplot(bd_avg, aes(x = time_sec, y = median_dir)) + 
  geom_path() + geom_point(aes(color=group)) +
  scale_x_continuous(limits = c(0,120), breaks = seq(0, 120, by = 10)) +
  labs(title = 'manual tracker, 1 frame/2 s')
```




```{r}
ggplot(filter(bd, time_sec<59|time_sec>61), aes(x = angle, y = dist)) +
         geom_point(alpha = 0.01) + geom_smooth(se=F)
```






```{r}
ggplot(filter(bdr, sec<58|sec>62)%>%sample_n(90000), aes(x = angle, y = distance)) + 
  geom_point(alpha = 0.1) +
  geom_smooth(se=F) +
  scale_y_continuous(limits = c(0, 30))
```







# Exploratory Analyses

Make graph comparing treatments.

```{r message=FALSE, warning=FALSE}
bd_avg <- group_by(bd, time_sec, trt)%>%
  summarize(dista = mean(dist, na.rm=T))

ggplot(bd_avg, aes(x = time_sec, y = dista, color = trt)) +
  geom_line()
```

The infected copepods are less active than the uninfected copepods. This is true both before and after the simulated attack. When we separate this plot into the different days of the experiment, we see a similar pattern. Maybe at the earliest time point, day 5, the infected copepods are less manipulated.

```{r}
bd_avg <- group_by(bd, time_sec, day, trt)%>%
  summarize(dista = mean(dist, na.rm=T))

ggplot(bd_avg, aes(x = time_sec, y = dista, color = trt)) +
  geom_line() +
  facet_wrap(~day)
```

Now, let's look at the pattern for just the uninfected copepods, separated by copepod family. They seem pretty similar, thought one family (III) might be a little more active in general.

```{r}
bd_avg <- filter(bd, trt != 'infected')%>%
  group_by(time_sec, cop_fam)%>%
  summarize(dista = mean(dist, na.rm=T))

ggplot(bd_avg, aes(x = time_sec, y = dista, color = cop_fam)) +
  geom_line() 
```

Now, separate the data by parasite family. The differences appear reasonably consistent, suggesting parasite family is more important than host family.

```{r}
bd_avg <- filter(bd, trt == 'infected')%>%
  group_by(time_sec, parasite_fam)%>%
  summarize(dista = mean(dist, na.rm=T))

ggplot(bd_avg, aes(x = time_sec, y = dista, color = parasite_fam)) +
  geom_line() 
```

Here is the combo of host and parasite genotypes.

```{r}
bd_avg <- filter(bd, trt == 'infected')%>%
  group_by(cop_fam, parasite_fam)%>%
  summarize(dista = mean(dist, na.rm=T))

ggplot(bd_avg, aes(x = parasite_fam, y = dista, color = cop_fam)) +
  geom_line(aes(group = cop_fam))
```

# Models

```{r}
library(lme4)

# null model
out0 <- lmer(dist ~ factor(day) + (1|fname) + (1|cop_name), filter(bd, trt == 'infected'))

# add cop_fam to null
outc <- lmer(dist ~ cop_fam + factor(day) + (1|fname) + (1|cop_name), filter(bd, trt == 'infected'))
anova(out0, outc, test = 'Chi')

# add parasite_fam to null
outp <- lmer(dist ~ parasite_fam + factor(day) + (1|fname) + (1|cop_name), filter(bd, trt == 'infected'))
anova(out0, outp, test = 'Chi')

# add cop x parasite to null
outcp <- lmer(dist ~ cop_fam * parasite_fam + factor(day) + (1|fname) + (1|cop_name), filter(bd, trt == 'infected'))
anova(outc, outcp, test = 'Chi')
anova(outp, outcp, test = 'Chi')
```

GxG!!! Awesome!!


```{r}
sort(table(bd$dist), decreasing = T)
qplot(data = filter(bd, dist > 0), x = dist, binwidth = 0.1) 
#+ coord_cartesian(xlim = c(0, 0.5))
```

